---
title: Effects of increased specialization on revenue of Alaskan salmon fishers over four decades
author: Eric J. Ward$^1$, Sean C. Anderson$^2$, Andrew O. Shelton$^1$, Richard E.
  Brenner$^3$, Milo D. Adkison$^4$, Anne H. Beaudreau$^3$, Jordan T. Watson$^5$, Jennifer
  C. Shriver$^3$, Alan C. Haynie$^6$, Benjamin C. Williams$^3$
date: ''
output:
  pdf_document:
    fig_caption: yes
    includes:
      in_header: options.sty
    latex_engine: xelatex
  html_document: default
  word_document: default
csl: journal-of-applied-ecology.csl
bibliography: CatchPortfolios_salmon.bib
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy=FALSE, tidy.opts=list(width.cutoff=60), warning = FALSE, message = FALSE)
library(knitr)
library(tidyverse)
library(ggsidekick)
library(ggrepel)
library(reshape2)
library(viridis)
library(date)
library(mgcv)
library(gridExtra)
library(zoo)
library(rstan)
library(ggrepel)
library(cowplot)
modelname = "model-linear-iid" # model-linear-nointeractions
```


$^1$Conservation Biology Division, Northwest Fisheries Science Center, National Marine Fisheries Service, National Oceanic and Atmospheric Administration, 2725 Montlake Blvd E, Seattle WA, 98112, USA\
$^2$School of Aquatic and Fishery Sciences, Box 355020, University of Washington, Seattle, WA 98195, USA\
$^3$Division of Commercial Fisheries, Alaska Department of Fish and Game, PO Box 115526, Juneau, AK 99811, USA\
$^4$College of Fisheries and Ocean Sciences, University of Alaska Fairbanks, 17101 Point Lena Loop Rd, Juneau, AK, 99801, USA\
$^5$Alaska Fisheries Science Center, National Marine Fisheries Service, National Oceanographic and Atmospheric Administration, Auke Bay Laboratories, 17109 Pt. Lena Loop Rd., Juneau, AK 99801, USA\
$^6$Alaska Fisheries Science Center, National Marine Fisheries Service, National Oceanographic and Atmospheric Administration, Bldg 4, 7600 Sand Point Way NE, Seattle, WA, 98115, USA\


## Abstract
\setlength\parindent{24pt}

1. Theory and previous studies have shown that commercial fishers with a diversify catch across multiple species may experience benefits 
 such as increased revenue and reduced variability in
revenue. However, fishers can only
increase the species diversity of their catch if they own fishing permits that allow
multiple species to be targeted, or if they own multiple single-species permits. Individuals holding a single permit can only increase catch diversity within the confines of their permit (e.g., by fishing longer or over a broader spatial area).
2. Using a large dataset of individual salmon fishers in
Alaska, we build a Bayesian variance-function regression model to understand
how diversification impacts revenue and revenue variability, and how these
effects have evolved since the 1970s.
3. Applying these models to six salmon fisheries that encompass a broad geographic range and a variety of harvesting methods and species, we find that the majority of these fisheries have experienced reduced catch diversity through time and increasing benefits of specialization on mean individual revenues, opposite of what theory predicts.
4. One factor that has been hypothesized to reduce catch diversity in salmon fisheries is large-scale hatchery production. While our results suggest negative correlations between hatchery returns and catch diversity for some fisheries, we find little evidence for a change in variability of annual catches associated with increased hatchery production.
5. We find that individuals participating in Alaska salmon fisheries do not always benefit from targeting a diverse catch portfolio. Fishers have some control over their own distribution of effort in space and time, but are also affected by a number of external factors including demand, prices offered by processors, and fluctuations in fish abundance. Life history variation of the species targeted may also play a role. Individuals participating in Alaskan fisheries with high contributions of pink salmon --- which have the shortest life cycles of all Pacific salmon --- also have the highest variability in year-to-year revenue.

## Introduction

Commercial fishers experience high degrees of variability in revenue [@sethi2012; @anderson_benefits]. Understanding the incentives and risks that fishers experience, as well as the diversification strategies they adopt to minimize year-to-year variability in revenue is important for maintaining sustainable livelihoods and productive fisheries. As in terrestrial agriculture [@purdy1997], diversifying catch may affect the mean or variance of revenue. Another parallel between fishers and farmers is that while the strategies of both may have some impact on revenue, a number of external drivers may also affect the ability or incentives to diversify. For commercial fishers, some of these factors include  the prices offered by processors, the relative abundance of species, and regulatory changes such as the adoption of permit or area restrictions. Because of the complexities of these factors, any benefits or costs of diversification may be variable through time.

Some of the most detailed analyses of relationships between catch diversification and revenue have been focused on Alaskan fisheries. Fisheries in Alaska are among the most productive in the world [@fissel2016] both by landed weight, and revenue  [@officeofscienceandtechnology2016] and individual fishers adopt a broad range of strategies intended to enhance profitability. Previous analyses have looked at effects of fishing strategies on communities [@sethi2014; @cline2017; @himes-cornell2015; @loring2016], vessels [@kasperski2013] and individual fishers [@anderson_benefits]. One emergent pattern across these scales is that diversity may act as a stabilizing force against catch variability. However, there are many instances where specialization may be beneficial, to the extent that individuals may accept the risks of higher variability when it is accompanied by higher mean revenue [@anderson_benefits].

Anderson et al. (2017) defined fishing strategies by the permit or collection of permits held by single individuals. Some permits in Alaskan waters allow only one species to be targeted (e.g., sablefish, halibut, herring, crab, shellfish, and sea cucumber), while other permits allow species groups to be targeted (e.g. groundfish, salmon). Individual fishers can diversify their catches by owning one of these multispecies permits, or by owning several single species permits. Understanding the effects of catch diversification on revenue of salmon fishers is important because though overall participation in Alaskan fisheries have declined since the 1980s, many Alaskan salmon fisheries have seen increases in participation since 2000 (Fig. A1). Many of the individuals who participate in salmon fisheries are increasingly salmon specialists, in that they only hold a single permit [@anderson_benefits].

Five species of salmon are harvested in Alaska waters (Chinook [*Oncorhynchus tshawytscha*], chum [*O. keta*], pink [*O. gorbuscha*], sockeye [*O. nerka*], and coho [*O. kisutch*]), and individual salmon fishers can diversify their catch among the five species. However, fishing permits are available for particular gear types by region within Alaska and in practice these fisheries show a range of species diversity (Fig. A2). Most salmon are caught with purse seines (6 permit-regions), drift gillnets (5), or set gillnets (12). While multiple salmon species are captured in some of these fisheries, there are also examples of fisheries where catch is dominated by a single species. Examples of specialized fisheries with low catch diversity include Bristol Bay set and drift gillnet fisheries that primarily harvest sockeye salmon. For these unique fisheries, specialization is associated with higher revenue and reduced variability in year-to-year revenue [@anderson_benefits]. Mechanisms hypothesized to influence specialization or diversification in salmon fisheries have been reviewed by Knapp (2012). These include a number of exogenous drivers, such as competition from farmed salmon production, consolidation of salmon buyers, marketing, prices, and new processing technology [@knapp_trends]. In the late 1980s and early 1990s, for example, salmon abundance was high and prices were low. Following price declines into the early 2000s [@adfg2007], some of these trends have more recently reversed (i.e., increasing prices and catches, Fig. 2; Figs. A3--A5). Advances in processing technology, such as freezing, may smooth prices across time because fish caught in one year may be sold the next [@knapp_trends].

Perhaps one of the largest changes affecting catch diversity or incentives for Alaskan salmon fishermen to diversify has been an increase in the production of hatchery salmon. Increased hatchery production may influence price or encounter rates and therefore alter the species available to particular gears, or alter incentives to target particular species. The modern era of Alaska's large-scale hatchery program began in the late 1970s in response to low returns [@stopha2016], with substantial increases in hatchery releases during the 1980s [@mahnken1998; @leber2008; @ward_evaluating]. Many of these hatchery programs increased production in an effort to stabilize the variability of salmon catches [@brooks1976; @hilborn2001a], and while recent work has suggested these programs may increase total yield [@amoroso2017], quantitative support for reduction in revenue variability remains unclear. Because many management regulations (season openers, mesh size) have been designed to specifically target hatchery salmon, we expect hatchery programs to affect catch diversity, though these effects have not been previously quantified. Previous work has largely concentrated on the ecological effects of hatcheries [@hilborn1992a; @hilborn2001a; @wertheimer2001; @ruggerone2015], or economic yield [@boyce1993; @leber2008]. Understanding the role salmon have in buffering or stabilizing fishing revenues is critical given the responses of salmon to variable ocean environments [@mueter2002; @schindler2010].

Here, we use a range of salmon fisheries in Alaska to examine the effect of catch diversification on revenue and revenue stability. We present summaries of the five main salmon gear types in each region across a wide range of salmon fisheries to understand the overall relationship among revenue, stability or variability of revenue, and diversity of catches. Focusing exclusively on salmon fisheries, we extend the dataset used by Anderson et al. (2017) to include 12 more years of data, 1975-2016. We also extend the time series models of Anderson et al. (2017) to include time-varying effects of catch diversification, to compare temporal trends across regions and to identify periods when having a diverse catch portfolio was most beneficial. Finally, we focus on a subset of salmon fisheries affected by hatcheries, examining how variation in hatchery production influences catch stability.

## Methods

### Data

We obtained fisheries landings and revenue data for all permit holders in Alaska from 1975-2016 from the Commercial Fisheries Entry Commission (CFEC). Reported gross earnings (revenue) associated with each fish ticket --- a mandatory record provided by processors of the weight and species of the catch delivered at each landing --- were adjusted for inflation by converting all revenues to 2009 USD [@unitedstatesbureauofeconomicanalysis2017]. Each fish ticket was associated with both a vessel and permit holder, and because of our focus on individuals, we aggregated data at the level of permit holders to generate annual summaries of catch diversity and revenue. Fisheries were defined as groups of individuals holding the same permit. While we use the entire CFEC dataset, subsets of data are useful to understand particular aspects of the relationship between revenue and diversity (e.g., fisheries with low species diversity, such as the Bristol Bay gillnet fisheries were not included in some models because individuals are highly specialized).

To identify shared and unique trends among salmon fisheries, summaries of revenue, revenue variability, and species diversity were calculated for all of the major salmon fisheries over the 41-year time-series. This includes information from permits in all of the five major gear types: purse seine, drift and set gillnets, hand troll, and power troll. In addition to calculating mean revenue, variability of revenue, and mean catch diversity for each fishery, we generated time series of each quantity to identify long term trends.

### Evaluation of effects of diversification over time for individual salmon fishers

To understand how altering catch diversification affects individual revenue and revenue variability, we focused on two gillnet and four purse seine salmon fisheries where effective diversity of the catch is larger relative to other salmon fisheries (Table 1; Figs. A6--A7). The data filtering steps described in Anderson et al. (2017) were applied to fish ticket data for years 1975--2016. We aggregated trip-level catch and revenue data to annual summaries for each individual permit holder. We restricted our analysis to only include permit holders who participated in a single fishery in a year; to avoid complications in modeling, pairs of years where individuals changed participation in fisheries from the first to the second year were not included.  Effective catch diversity was used as a proxy for diversification, calculated as the inverse of Simpson's diversity [@jost2006a; @anderson_benefits; @kasperski2013], with each species' contribution determined in terms of revenue, rather than landed weight. Effective catch diversity values near 1.0 indicate that an individual's revenue is dominated by a single species, whereas an effective catch diversity of 3.0 indicates that three species contribute equally to revenue. Briefly, we implemented a Bayesian time series model with group-level predictors and individual covariates (diversity of catch, days fished) to simultaneously estimate the mean and variance of revenue for individual fishers [@anderson_benefits]. 

Following Anderson et al. (2017), we used a class of regression models to simultaneously estimate the effects of catch diversification on an individuals' mean annual revenue and variability in annual revenue. The natural log of revenue for individual *i* in year *t* is modeled as $${ \ln  \left( { R }_{ i,t } \right)  } ={ B }_{ 0,j,t }+{ B }_{ 1,j,t }\cdot \Delta S_{ i,t }+{ B }_{ 2,j,t }\cdot \left( \Delta S_{ i,t }\cdot { S }_{ i,t-1 } \right) + { B }_{ 3,j }\cdot \Delta D_{ i,t }+\ln  \left( { R }_{ i,t-1 } \right)$$ where *j* subscripts the salmon permit (Table 1). The variables $\Delta { S }_{ t }=\ln\left( { S }_{ t }/{ S }_{ t-1 } \right)$ represent the percent change in species diversity of the catch and $\Delta { D }_{ t }=\ln\left( { D }_{ t }/{ D }_{ t-1 } \right)$ represents the percent change in fishing effort (days fished) between years. Time-varying intercepts, specific to each permit (Table 1), were modeled as random and exchangeable across years, ${ B }_{ 0,j,t }\sim \mathrm{Normal}\left( { u }_{ { B }_{ 0 } },{ \sigma  }_{ { B }_{ 0 } } \right)$. The term ${ B }_{ 1,j,t }$ represents the permit-level random effect of increasing catch diversity. We also included the terms $\Delta S_{ t }\cdot { S }_{ t-1 }$, which forced the effect of diversification to be linear but allowed its strength to vary through time. The coefficients ${ B }_{ 1,j,t }$ and ${ B }_{ 2,j,t }$ were included to allow the interaction to vary by permit and year. The flexibility of this model allowed the benefit of a fisher increasing catch diversity to be positive in some years (generalizing translates to increased revenue), negative in other years (specializing increases revenue) and neutral in other years (revenue is independent of alterations to catch composition). We constrained the random effects associated with diversification to have shared variances across strategies (permits), for example: ${ B }_{ 1,j,t }\sim \mathrm{Normal}\left( 0, { \sigma  }_{ { B }_{ 1 } } \right)$. The effect of changing effort (measured as percent change in days fished) ${ B }_{ 3,j }$ was allowed to vary by strategy, but we did not allow this effect to vary through time. Finally, the term $\ln\left( { R }_{ i,t-1 } \right)$ was included as an offset because individual revenues are typically non-stationary.

The second component of this regression model involved effects on the variability of annual revenue.  Modeling the mean and variance simultaneously makes the model akin to variance function regression or stochastic volatility models [@smyth1989; @western2009; @anderson_benefits]. Our model for the standard deviation was $$\ln\left( { \sigma  }_{ i,t } \right)={ G }_{ 0,j }+{ G }_{ 1,j,t }\cdot \Delta S_{ i,t }+{ G }_{ 2,j,t }\cdot \left( \Delta S_{ i,t }\cdot { S }_{ i,t-1 } \right) +G_{ 3,j }\cdot \Delta D_{ i,t }.$$ We constrained the variance intercepts ${G}_{0,j}$ to be constant by permit. As in with the mean model, the random effects were allowed to be independent by permit and year, ${ G }_{ 1,j,t }\sim \mathrm{Normal}\left( 0, { \sigma  }_{ { G }_{ 1 } } \right)$, and covariates associated with days fished, $\Delta D_{ i,t }$, and species diversity $\Delta S_{ i,t }$ were included as predictors.

### Relationships between hatchery production, diversity and catches

To quantify potential effects of hatchery production on diversity and catch variability we used data from the seine and gillnet fisheries in Prince William Sound (PWS) and Southeast Alaska (SE) and assessed correlations in catch diversity (Figs. A6--A7) with returns of dominant hatchery species (Fig. 1). In Southeast Alaska, we examined correlations in catch diversity with hatchery chum returns, because chum are caught by both seine and gillnet fisheries (Figs. A6--A7). For PWS, we calculated the correlations of catch diversity of the seine fishery with hatchery returns of pink salmon (Fig. A6) and the catch diversity of the gillnet fishery with hatchery returns of sockeye salmon (Fig. A7).

To quantify relationships between hatchery production and catch variability, we used the landed catch from each region and species as a response [@amoroso2017; @hilborn2001a; @knapp2007]. Using a variance function regression model, we extended the approach of [@amoroso2017] to model variability in salmon catches as a continuous function of the hatchery contribution of catches. Within each region (Table 1), the hatchery contribution to catches and total catches of each species are from Stopha (2016), with Kodiak and the Alaska Peninsula grouped because of reporting. The catch of each species in each region was modeled as, $$E[{ Y }_{ s,f,t }]={ Y }_{ s,f,t-1 }+{ b }_{ s,f }{ X }_{ s,f,t }$$ where ${ Y }_{ s,f,t-1 }$ is the catch of species $s$ in region $f$ in the previous time step $t-1$, ${ b }_{ s,f }$ is a coefficient relating the contribution of hatchery fish to the mean catch, and ${ X }_{ s,f,t }$ is the change in the proportion of hatchery contribution to the total catch for species *s* between times *t* and *t-1*. The hatchery-effect coefficients were modeled hierarchically, so ${ b }_{ s,f }\sim \mathrm{Normal}\left( { u }_{ b },{ \sigma  }_{ b } \right)$, where ${ u }_{ b }$ and ${ \sigma  }_{ b }$ represent the mean and standard deviation of the random effects. Our second equation relates the change in proportion of hatchery contribution to the catch variability, $$\ln\left( { \sigma  }_{ s,f,t } \right) ={ g  }_{ 0,s,f }+{ g  }_{ 1,s,f }{ X }_{ s,f,t }$$ where ${ g  }_{ 0,s,f }$ represents an intercept unique to each region and species, and ${ g  }_{ 1,s,f }$ represents a region- and species-specific slope, ${ g }_{ 1,s,f }\sim \mathrm{Normal}\left( { u }_{ g },{ \sigma  }_{ g } \right)$. Combining the regression of the mean and variance, the observed catches were assumed to be normal $\ln\left( { Y }_{ s,f,t } \right) \sim \mathrm{Normal}\left( E\left[ \ln\left( { Y }_{ s,f,t } \right)  \right] ,{ \sigma  }_{ s,f,t } \right)$.

### Estimation

Estimation for both sets of models was done in a Bayesian framework using R [@rcoreteam2017] and the package `rstan` [@standevelopmentteam2016], implementing Markov chain Monte Carlo (MCMC) using the No-U Turn Sampling (NUTS) algorithm [@hoffman2014; @standevelopmentteam2015; @standevelopmentteam2015a]. Following Anderson et al. (2017), each model included three MCMC chains, with a warmup period of 4000 samples, followed by 5000 saved iterations. Posterior estimates were visually examined for chain convergence with trace plots, and the potential scale reduction factor [@gelman1992a] was used as a tool to ensure convergence ($\hat { R }$ < 1.05). Code to replicate our analysis, including additional diagnostics plots, and additional models is available at <https://github.com/NCEAS/pfx-commercial-salmon>.

## Results

Since the early 2000s, Alaskan salmon fisheries have generally experienced an increase in individual participants, landings have burgeoned faster than revenue, and species diversity of landed catch has declined over time (Fig. 2; Figs. A1, A6-A7). Across salmon fisheries, there appears to be little correlation between catch diversity and revenue (Fig. A2), though fisheries cluster by gear type on axes of risk (CV of annual revenue) and return (mean annual revenue, Fig. 3).

Most fishers participating in the six salmon fisheries in our analysis experienced changes in effective catch diversity that are typically increase or decrease by one species per year (Fig. A8). Our model of individual revenues suggests that catch diversity in these six fisheries has declined over time (Figs. A6--A7). The estimated negative effects of increasing diversity on mean revenue (Fig. 4) implies that for the majority of these fisheries in most years, revenue increases when individuals specialize and reduce the diversity of their catches. Across fisheries and years, the average range (95% CIs) suggests that for an individual earning 100000 dollars, increasing effective catch diversity from one to two species would be expected to reduce revenue to 68000 dollars or increase revenue up to 130000 dollars. The slight negative trends in these estimated effects for some fisheries in recent years (PWS and SE drift gillnet) suggest that the benefits of specializing have also increased over time (Fig. 4). More importantly, our estimates suggest that individuals experience changing incentives to diversify over time. From 2008--2010 for example, it was not beneficial for individuals in the PWS or SE seine fisheries to diversify (negative effects in Fig. 4), but several years later these trends reversed for seine fisheries, indicating a more diverse catch translated into higher revenue (Fig. 4).

Our results also showed contrasts in year-to-year variability between seine and gillnet fisheries, with individuals in seine fisheries experiencing higher annual variability in revenue (Fig. A9). Across individual fishers, seine fishers are generally higher on both axes of risk and return (CV of annual revenue versus mean annual revenue; Fig. 3). Though the catch composition has fluctuated through time in areas where seine and gillnet fisheries exist (PWS, SE), catch diversity is slightly higher in gillnet fisheries (Figs A6--A7). In most fisheries, we found little evidence that individuals who increased catch diversity reduced their revenue variability (Fig. A10). For several fisheries (SE seine and gillnet, PWS gillnet), there have been periods over the last several decades where diversification has increased variability, such as 2008--2010 (Fig. A10). For most other fisheries, these effects of diversifying are near zero (Fig. A10).

For the four fisheries in PWS and SE that are influenced by hatchery production, we found negative correlations between the share of the dominant hatchery species and total catch diversity in all but the SE seine fishery (Fig. 5). These correlations were strongest for SE gillnet fisheries targeting chum salmon $\rho = -0.76$ ($P < 0.01$) and for the PWS fishery targeting pink salmon $\rho = -0.41$ ($P < 0.01$).

We found no consistent support for strong positive or negative relationships between variability of catches and the hatchery contribution to returns based on model results. To allow for comparison across regions and scales of data, we converted coefficient estimates to effect sizes, calculating the approximate percent change in standard deviation resulting from the hatchery contribution increasing by 20%, ${ \kappa  }_{ s }=\ln\left( \frac { { g }_{ 0,s }+{ 0.2 \cdot g }_{ 1,s } }{ { g }_{ 0,s } }  \right)$. For the three regions in our analyses, most estimates of these effect sizes are near zero (Fig. 6). The largest positive correlations between hatchery contribution and variability in catches appeared to be for coho and pink salmon in PWS (Fig. 6), and a slight negative effect was found between the hatchery contributions and Chinook salmon catches in SE. The 95% credible intervals for these estimates overlap zero, however, and the estimated effect sizes are relatively small in magnitude. In other words, a 20% increase in hatchery fish results in less than a 20% increase in the standard deviation of catches.

## Discussion

Determining the mechanisms responsible for trends or variability in the revenue of individuals whose incomes are dependent on natural resources is important for understanding the dynamics of coupled social-ecological systems. Income variability for such individuals is typically greater than that of the larger population [@mishra2002; @sethi2012]. Individuals participating in some commercial salmon fisheries in Alaska may have the option to increase the diversity of their catches. Diversification strategies may include buying additional permits, but many fishers have become single-permit holders [@anderson_benefits].  For those individuals, targeting additional species allowed by the permit is the only option for diversifying catch. Our analysis illustrates that catch diversity has declined in some salmon fisheries. In recent years, catches from several of these fisheries have been dominated by a single species, namely sockeye salmon in the Prince William Sound gillnet fishery, pink salmon in the Prince William Sound seine fishery, and chum salmon in the Southeast Alaska gillnet fishery (Figs. A6--A7).

As in other commercial fisheries, the ability of salmon fishers to diversify within the constraints of their permits may be largely driven by factors beyond their control. The limited options of salmon fishers in Alaska to diversify may include fishing in different months or different regions. For example, individuals participating in the Prince William Sound seine fishery may fish earlier in the year to target chum salmon. External incentives for salmon fishers to specialize or diversify are complex, and include a suite of economic factors: competition from foreign markets, accessibility to buyers and prices from processors, fuel costs, and marketing (e.g., increased marketing of Copper River sockeye salmon beginning in the early 1980s; Knapp 2012). A potential driver unique to Alaska salmon fisheries is the potential effects of hatchery production. Increased hatchery production may alter the species composition that individual fishers encounter, particularly if fishing seasons are designed to intercept large hatchery runs. Prices of all salmon have generally increased since the early 2000s (Figs. A4--A5), with some of the largest changes being in pink and chum salmon prices. In areas with high pink or chum salmon runs, the benefits of diversification may have been greatest when these species had low returns. For pink salmon returns with strong 2-year oscillations for example, this would correspond to even- versus odd-years.

Results from our regression model of individual fisher revenues over the period 1975--2016 suggest that there have been long-term increases in the benefits of individual fishers specializing (i.e., expected revenue increases when catch diversity decreases; Fig. 5). For most of the time series, the impacts of changes to catch diversity on variability in year-to-year revenue are less clear. However, a counter-intuitive result is that since the mid-2000s, individuals participating in fisheries in Prince William Sound and seine fisheries in Southeast Alaska experienced an increase in revenue variability associated with increased diversification (Fig. A10). Some of this relationship may be spurious, in that individuals who reduce effort and catches because of unmodeled factors (e.g. illness, gear or vessel malfunctions) and this reduction in landings may be accompanied by a relative increase in species diversity (Fig. A11). 

An additional potential effect of increased hatchery production may be changes in the variability of catches and/or revenues to fisheries or individuals. Across multiple regions in Alaska, we found little support for catches stabilizing as hatchery programs have increased. However, one example of a positive correlation between the hatchery contribution to harvest variability is pink salmon in Prince William Sound (Fig. 6). Our model describing the revenues of individual fishers also suggests that individuals in the Prince William Sound seine fishery have higher revenue variability relative to other fisheries (Fig. 2–3, A9), with year-to-year revenue changes being correlated with fluctuations in pink salmon returns (Fig. A12). Pink salmon have the least variable life-history characteristices of all Pacific salmon, with a fixed 2-year life cycle and well-documented odd or even year dominance for a particular area [@quinn2005a]. The catch composition of the PWS seine fishery is heavily dominated by pink salmon and has a low level of species diversity (Fig. 1–2, 4). Therefore, revenue variability for this fishery may be a function of the life history traits of pink salmon (which are mostly of hatchery origin). In contrast, the SEAK seine fishery has a low level of catch diversity but is dominated by wild pink salmon from approximately 2,000 streams [@dangel1988] and large numbers of hatchery chum salmon, and may be buffered against fluctuations in revenue (Figs. 2--3) by the portfolio of multiple populations [e.g., @schindler2010] and—for chum salmon—a more diverse suite of life history characteristics.

The increased specialization of Alaska fisheries has been highlighted in several recent analyses [@sethi2012; @kasperski2013; @anderson_benefits], and our analysis extends this work to show similar trends within individual salmon permits. In terrestrial agriculture, there are multiple examples of year-to-year income variability being stabilized by diversification [@difalco2003], but these incentives become more complex at larger scales. Crop diversification in developed countries is shaped by a number of drivers including changing national policies, technology, and increased demand from other parts of the world (e.g., U.S. production of soybeans for Asian countries) [@bradshaw2004; @aguilar2015]. Shifts in terrestrial agriculture from a diverse mix of low- and high-value crops to focusing on high-volume low-value crops (such as soybeans or cereals) may have parallels for Alaskan salmon fisheries --- particularly those fisheries that have more recently targeted low-value species that faster life history traits (pink and chum salmon). 

Understanding patterns of variability that individuals experience is essential for quantifying changing revenue patterns and risk at the level of fisheries and coastal communities. However, protecting individuals from future economic shocks is complicated, and may be affected by factors including the relative composition of species, the spatial distribution of fish, market forces, hatchery programs, or environmental variation. Management tools to counter risk are generally limited, and vary by country. In the USA, crop insurance programs used by terrestrial farmers have been considered but not implemented [@herrmann2004]. Losses may also be offset through federally funded fishery disaster assistance programs through the Magnusson Stevens Act. Fishery disasters have been declared in Alaska salmon fisheries in 1997-2000 (80 million USD), 2009-2012 (> 25 million USD) and for pink salmon in 2016 (amount to be determined). As some of these salmon fisheries have become more specialized, management may benefit from future work into how specialization affects the likelihood of disaster declaration, and how disaster funding affects individual fishers. 

## Acknowledgments
This project was conducted as part of a NCEAS working group, "Applying Portfolio Effects to the Gulf of Alaska Ecosystem" (led by K. Marshall). Data sets included in our analysis have been provided by agencies (ADF&G, NOAA) and have been made available on the EVOS Data Portal, http://gulfwatch.nceas.ucsb.edu/#data/page/0. ADF&G retains intellectual property rights to data collected by or for ADF&G. Any dissemination of the data must credit ADF&G as the source, with a disclaimer that exonerates the department for errors or deficiencies in reproduction, subsequent analysis, or interpretation. The authors thank staff at NCEAS, and additional post-docs for support (R. Blake, C. Ward). A. Munro, C. Siddon, and B. Templin for providing helpful reviews that greatly improved this manuscript.

## Author Contribution Statement
EW, AB, RB, and OS secured funding for this project. CFEC data was provided by JS and JW with assistance from RB. EW and SA did the modeling, with EW, SA, OS, and JW producing the main and supplemental figures, with guidance from JS and RB. All authors contributed to performing preliminary analyses as part of the NCEAS workshops, critically assisted in writing drafts and gave final approval for publication.

\break

```{r fig1, echo=FALSE, fig.pos="placeHere", fig.cap="Harvest of hatchery-produced salmon through time by species and region. Shown are the numbers of hatchery fish caught by fisheries (top row), and the hatchery contribution of each species to total catch (bottom row). Regions represents combined Kodiak and Alaska Peninsula (KO - AP), Southeast (SE) and Prince William Sound (PWS). "}

seak_hatch = read.csv("../data-generated/seak_hatcheryHarvest.csv")
pws_hatch = read.csv("../data-generated/pws_hatcheryHarvest.csv")
kodiak_hatch = read.csv("../data-generated/kodiak_hatcheryHarvest.csv")

seak_hatch = melt(seak_hatch, id.vars = c("Year"))
pws_hatch = melt(pws_hatch, id.vars = c("Year"))
kodiak_hatch = melt(kodiak_hatch, id.vars = c("Year"))

seak_hatch$region = "SE"
pws_hatch$region = "PWS"
kodiak_hatch$region = "KO - AP"

seak_hatch$value[is.na(seak_hatch$value)] = 0
pws_hatch$value[is.na(pws_hatch$value)] = 0
kodiak_hatch$value[is.na(kodiak_hatch$value)] = 0

dat = bind_rows(seak_hatch, pws_hatch)
dat = bind_rows(dat, kodiak_hatch)

names(dat)[which(names(dat)=="variable")]="species"
dat$value = dat$value / 1000000

# chinook chum coho pink sock
cbPalette <- c("#009E73", "#4B2E83", "#999999", "#CC79A7", "#D55E00")

dat$col = as.numeric(as.factor(dat$species))
g1 = ggplot(dat, aes(Year, value)) +
  geom_area(aes(fill = species), position = "stack") +
  xlab("Year") + ylab("Hatchery harvest (millions)") +
  scale_fill_manual(values = c('Chinook'=cbPalette[1], 'Chum'=cbPalette[2],
    'Coho'=cbPalette[3], 'Pink'=cbPalette[4], 'Sockeye'=cbPalette[5]), name="Species",
    labels =c("Chinook","Sockeye","Coho","Pink","Chum")) +
  facet_wrap(~region, scale="free_y") + theme_sleek()

# bring in total harvest data (hatchery + wild)
hatch = read.csv("../data-generated/commonPropertyCommercialCatch.csv",
  stringsAsFactors = FALSE)

kodap1 = hatch[which(hatch$Region%in%c("Kodiak") & hatch$Year <= 1995),]
kodap2 = hatch[which(hatch$Region%in%c("AP", "Kodiak") & hatch$Year > 1995),]
kodap2 = dplyr::group_by(kodap2, Year) %>%
  summarize(Chinook_harvest = sum(Chinook_harvest),
    Sockeye_harvest = sum(Sockeye_harvest),
    Coho_harvest = sum(Coho_harvest),
    Pink_harvest = sum(Pink_harvest),
    Chum_harvest = sum(Chum_harvest),
    Region = Region[1],
    Source = Source[1],
    ReportNumber = ReportNumber[1],
    Chinook_enhanced = sum(Chinook_enhanced),
    Sockeye_enhanced = sum(Sockeye_enhanced),
    Coho_enhanced = sum(Coho_enhanced),
    Pink_enhanced = sum(Pink_enhanced),
    Chum_enhanced = sum(Chum_enhanced))
kodap = rbind(as.data.frame(kodap1), as.data.frame(kodap2))
kodap$Region = as.character(kodap$Region)
kodap$Source = as.character(kodap$Source)
kodap$ReportNumber = as.character(kodap$ReportNumber)

hatch = rbind(hatch[which(hatch$Region %in% c("PWS","Southeast")),], kodap)

hatch$pChinook = hatch$Chinook_enhanced/hatch$Chinook_harvest
hatch$pSockeye = hatch$Sockeye_enhanced/hatch$Sockeye_harvest
hatch$pCoho = hatch$Coho_enhanced/hatch$Coho_harvest
hatch$pPink = hatch$Pink_enhanced/hatch$Pink_harvest
hatch$pChum = hatch$Chum_enhanced/hatch$Chum_harvest

hatch = hatch[,c("Year","pChinook","pSockeye","pCoho","pPink","pChum","Region")]
hatch = dplyr::rename(hatch, Chinook=pChinook,Sockeye=pSockeye, Coho=pCoho, Pink=pPink, Chum=pChum)
hatch$Region = as.character(hatch$Region)
hatch$Region[which(hatch$Region=="PWS")] = "PWS"
hatch$Region[which(hatch$Region=="Kodiak")] = "KO - AP"
hatch$Region[which(hatch$Region=="Southeast")] = "SE"
hatch = melt(hatch, id.vars = c("Year","Region"))
names(hatch)[which(names(hatch)=="variable")]="Species"

g2 = ggplot(hatch, aes(Year, value, color=Species)) +
  geom_line() + facet_wrap(~Region) +
  xlab("Year") + ylab("Hatchery fraction of total catch") +
  scale_color_manual(values = c('Chinook'=cbPalette[1], 'Chum'=cbPalette[2],
    'Coho'=cbPalette[3], 'Pink'=cbPalette[4], 'Sockeye'=cbPalette[5]), name="Species",
    labels =c("Chinook","Sockeye","Coho","Pink","Chum")) + theme_sleek()

gridExtra::grid.arrange(g1, g2)

# join datasets together for modeling
names(dat) = tolower(names(dat))
names(hatch) = tolower(names(hatch))
names(dat)[which(names(dat)=="value")] = "hatchery_harvest_millions"
names(hatch)[which(names(hatch)=="value")] = "percent_hatchery"
dat = left_join(dat, hatch)
dat$total_harvest = dat$hatchery_harvest_millions / (dat$percent_hatchery/100)
saveRDS(dat, file="../data-generated/hatchery_harvest_millions.rds")
```

\break

```{r fig2, echo=FALSE, fig.pos="placeHere", fig.cap="Trends in catch diversity, landed catch (metric tons), participation (permit holders), and mean annual revenue for eight selected Alaska salmon fisheries. The troll fishery permits (hand and power troll) do not have an associated area (statewide) but the majority of fishing is done in Southeast Alaska. Areas associated with permits include the Alaska Peninsula (AP), Bristol Bay (BB), Kodiak (KO), Prince William Sound (PWS), Southeast Alaska (SE)."}
# trends in participation,
cfec_all = readRDS("../data-generated/salmon_data_for_paper.rds")
people = group_by(cfec_all, p_fshy, year) %>%
  summarize(n = length(unique(p_holder))) %>%
  filter(substr(p_fshy,1,1)=="S")
people$p_fshy = paste0(substr(people$p_fshy,1,1), substr(people$p_fshy,3,5))
names(people)[which(names(people)=="p_fshy")]="Permit"

people2 = people[people$Permit%in%c("S01K","S01M","S01E","S01A","S03A","S03E","S03T","S05B"),]
people2$Permit[which(people2$Permit=="S01K")] = "KO-seine"
people2$Permit[which(people2$Permit=="S01M")] = "AP-seine"
people2$Permit[which(people2$Permit=="S01E")] = "PWS-seine"
people2$Permit[which(people2$Permit=="S01A")] = "SE-seine"
people2$Permit[which(people2$Permit=="S03A")] = "SE-drift"
people2$Permit[which(people2$Permit=="S03E")] = "PWS-drift"
people2$Permit[which(people2$Permit=="S03T")] = "BB-drift"
people2$Permit[which(people2$Permit=="S05B")] = "State-troll"

people2 = rename(people2, permit = Permit)
people2 = rename(people2, value = n)
people2$value = log10(people2$value)
people2$datatype = "log10 (permit holders)"

# trends in landed weight - metric tons
weight = group_by(cfec_all, p_fshy, year) %>%
  summarize(w = sum(g_pounds,na.rm=T) * 0.000453592) %>%
  filter(substr(p_fshy,1,1)=="S")
weight$p_fshy = paste0(substr(weight$p_fshy,1,1), substr(weight$p_fshy,3,5))
names(weight)[which(names(weight)=="p_fshy")]="Permit"

weight = left_join(as.data.frame(weight), as.data.frame(people))

weight2 = weight[weight$Permit%in%c("S01K","S01M","S01E","S01A","S03A","S03E","S03T","S05B"),]
weight2$Permit[which(weight2$Permit=="S01K")] = "KO-seine"
weight2$Permit[which(weight2$Permit=="S01M")] = "AP-seine"
weight2$Permit[which(weight2$Permit=="S01E")] = "PWS-seine"
weight2$Permit[which(weight2$Permit=="S01A")] = "SE-seine"
weight2$Permit[which(weight2$Permit=="S03A")] = "SE-drift"
weight2$Permit[which(weight2$Permit=="S03E")] = "PWS-drift"
weight2$Permit[which(weight2$Permit=="S03T")] = "BB-drift"
weight2$Permit[which(weight2$Permit=="S05B")] = "State-troll"

weight2$wn = log10(weight2$w/weight2$n)
weight2 = rename(weight2, permit=Permit, value = wn)
weight2$datatype = "log10 (avgerage catch)"
datf = as.data.frame(rbind(people2[,c("permit","year","value","datatype")],
  weight2[,c("permit","year","value","datatype")]))

# trends in revenue
dat.annual = readRDS(file="../data-generated/salmon-annual-for-modeling.rds")
dat.annual = group_by(dat.annual[which(dat.annual$strategy_permit %in% c("S01A", "S01E", "S01K", "S01M", "S03A", "S03E", "S03T", "S05B")),], year, strategy_permit) %>%
  summarize(meanrev = mean(revenue), meandiv = mean(specDiv))
names(dat.annual)[which(names(dat.annual)=="strategy_permit")]="Permit"
dat.annual = left_join(dat.annual, weight)

dat.annual2 = dat.annual[dat.annual$Permit%in%c("S01K","S01M","S01E","S01A","S03A","S03E","S03T","S05B"),]
dat.annual2$Permit[which(dat.annual2$Permit=="S01K")] = "KO-seine"
dat.annual2$Permit[which(dat.annual2$Permit=="S01M")] = "AP-seine"
dat.annual2$Permit[which(dat.annual2$Permit=="S01E")] = "PWS-seine"
dat.annual2$Permit[which(dat.annual2$Permit=="S01A")] = "SE-seine"
dat.annual2$Permit[which(dat.annual2$Permit=="S03A")] = "SE-drift"
dat.annual2$Permit[which(dat.annual2$Permit=="S03E")] = "PWS-drift"
dat.annual2$Permit[which(dat.annual2$Permit=="S03T")] = "BB-drift"
dat.annual2$Permit[which(dat.annual2$Permit=="S05B")] = "State-troll"

dat.annual2 = rename(dat.annual2, permit=Permit)

dat.annual2$meanrev = log10(dat.annual2$meanrev)
tempd = as.data.frame(dat.annual2[,c("permit","year","meanrev")])
names(tempd) = c("permit","year","value")
tempd$datatype = "log10 (revenue)"
datf = rbind(datf, tempd)

tempd = as.data.frame(dat.annual2[,c("permit","year","meandiv")])
names(tempd) = c("permit","year","value")
tempd$datatype = "Catch diversity"
datf = rbind(datf, tempd)

datf = rename(datf, Permit = permit)

datf$value[which(datf$year==1989 & datf$Permit == "KO-seine")] = NA

ggplot(datf, aes(year, value, color = Permit, group = Permit)) + geom_line() + ylab("") + xlab("Year") + facet_wrap(~datatype, scale="free_y") + theme_sleek()

```

\break

```{r fig3, echo=FALSE, fig.pos="placeHere", fig.cap="Coefficient of variation (CV) and mean annual revenue by salmon permit type, 1975--2016. Effective species diversity of catch (1 / Simpson's diversity) is represented by circle area, fishing gear types are shown with color shading. Labels indicate regions --- Alaska Peninsula (AP), Bristol Bay (BB), Chignik (CH), Cook Inlet (CI), Kodiak (KO), Kuskokwim (KU), Norton Sound (NS), Prince William Sound (PWS), Southeast Alaska (SE), Yakutat (YAK). The troll fishery permit does not have an associated area (statewide) but the majority of fishing is done in Southeast Alaska.", warning=FALSE, message=FALSE}

dat.annual <- read.table("../data-generated/bubble_plot_data_fig2.csv", stringsAsFactors = FALSE, header = TRUE)

dat.annual <- mutate(dat.annual,
  gear = factor(gear, levels = c("Drift gillnet", "Set gillnet", "Hand troll", "Power troll", "Purse seine")))

l <- tibble::tribble(~let, ~label,
  "H", "CI",
  "E", "PWS",
  "L", "CH",
  "K", "KO",
  "M", "AP",
  "T", "BB",
  "A", "SE",
  "B", "State",
  "Z", "NS",
  "W", "KU",
  "D", "YAK",
  "Y", "YUK")

dat.annual <- mutate(dat.annual,
  let = sub("S[0-9]+([A-Z])", "\\1", dat.annual$strategy_permit)) %>%
  left_join(l) %>%
  filter(strategy_permit != "S04Y")

g <- ggplot(dat.annual,
  aes(x = mean_rev/1e3, y = cv_rev,
    label = label)) +
  geom_point(alpha = 0.5, aes(fill = gear, size = diversity), pch = 21,
    colour = "grey20") +
  scale_size(range = c(1.5, 15), breaks = seq(1.2, 2.4, 0.4)) +
  guides(fill = guide_legend(override.aes = list(size = 4, alpha = 0.7), order = 1)) +
  geom_text_repel(size = 2.6, colour = "grey20",
    alpha = 1, segment.color = "grey70",
    point.padding = unit(0, "lines"), max.iter = 1e4, segment.size = 0.3) +
  scale_fill_brewer(palette = "Set3") +
  ylab("CV annual revenue") + xlab("Mean revenue ($1000)") +
  labs(fill = "Gear", size = "Diversity") +
  scale_y_continuous(breaks = seq(0.2, 1, 0.2)) +
  theme_sleek() #+
  # scale_x_log10()
print(g)
```

\break

```{r fig4, fig.pos="placeHere", fig.cap="Time-varying effects of increasing the effective species diversity of catch (1/Simpson's diversity) from 1 to 2 species on mean revenue (presented as log ratios). The solid lines indicate the mean effect and the shaded region represents 95% CIs. Labels indicate regions and gear type --- Alaska Peninsula (AP), Kodiak (KO), Prince William Sound (PWS), Southeast Alaska (SE) seine net or drift gill net.", echo=FALSE}
# make yearpermit df
diffdat = readRDS(file="../data-generated/salmon.rds")
years = seq(min(diffdat$year), max(diffdat$year))
permits = levels(as.factor(diffdat$permit))
# note that we have to use factor() here instead of unique() because
# that's the same as the ordering in the estimation code
yearpermit = data.frame("yrpermit"=levels(as.factor(diffdat$year_permit)))
yearpermit$year = as.numeric(substr(yearpermit$yrpermit,1,4))
yearpermit$permit = substr(yearpermit$yrpermit,6,9)

# grab original data to look at permit names
diffdat = readRDS(file="../data-generated/salmon-diff-for-modeling.rds")
years = seq(min(diffdat$year), max(diffdat$year))
#permits = levels(as.factor(diffdat$permit))

yearpermitold=yearpermit
load("../analysis/model-linear-iid.Rdata")
yearpermit=yearpermitold

est = extract(mod, permute="TRUE")

# start with making time series of estimates on mean
# challenge is that they're time varying quadratic functions
# do facet_wrap of estimates by year

# effect of adding more 1 species from 1. specdiv.prev = 1

specdiv.prev = seq(1, 5, by=0.1)
specdiv = specdiv.prev + 1 # always add 1 more in these comparisons
x1 = log(specdiv/specdiv.prev)
x2 = log(specdiv/specdiv.prev) * specdiv.prev
x3 = log(specdiv/specdiv.prev) * specdiv.prev^2

df = expand.grid("year"=1:length(years), "diversity" = seq(1,length(x1)), "permit"=1:6)
df$permits = permits[df$permit]
df$years = years[df$year]
df$x = specdiv.prev[df$diversity]
df$pred = 0
df$pred_median = 0
df$low = 0
df$high = 0
b0_indx = match(paste(as.character(df$years), as.character(df$permits)), as.character(yearpermit$yrpermit))
for(i in 1:nrow(df)) {
  #tmp = est$b0_str_yr[,b0_indx[i]] +
  tmp =
    x1[df$diversity[i]] * (est$b_1[,df$year[i], df$permit[i]])# +
    #x2[df$diversity[i]] * (est$b_2[,df$year[i], df$permit[i]]+est$b2_str_yr_mu)# +
    #x3[df$diversity[i]] * (est$b_3[,df$year[i], df$permit[i]]+est$b3_str_yr_mu)
  df$pred[i] = mean(tmp)
  df$pred_median[i] = median(tmp)
  df$low[i] = quantile(tmp,0.025)
  df$high[i] = quantile(tmp,0.975)
}

df$permits[which(df$permits=="S01A")] = "SE-seine"
df$permits[which(df$permits=="S01E")] = "PWS-seine"
df$permits[which(df$permits=="S01K")] = "KO-seine"
df$permits[which(df$permits=="S01M")] = "AP-seine"
df$permits[which(df$permits=="S03A")] = "SE-drift"
df$permits[which(df$permits=="S03E")] = "PWS-drift"

ggplot(df[df$diversity==1,], aes(x=years,y=pred,group=permits)) +
  geom_ribbon(aes(ymin=low,ymax=high), fill="blue",alpha=0.4) +
  geom_line(col="blue") +
  facet_wrap(~permits) + xlab("Year") + ylab("Effect of increased catch diversity on mean revenue") + geom_hline(yintercept=0, color="red", linetype = 2, alpha = 0.3) + theme_sleek()

# for revision,
#group_by(df, permits) %>% summarize(lo = min(low), hi = max(high)) # %>% ungroup() %>% summarize(lo = mean(lo), hi = mean(hi))
```

\break

```{r fig5, echo=FALSE, fig.pos="placeHere", fig.cap="Catch diversity (1 / Simpson's diversity) and harvest of dominant hatchery species by year. For Southeast Alaska gillnet and seine fisheries we used harvest of hatchery chum salmon, for the PWS gillnet fishery we used harvest of hatchery sockeye salmon, and for the PWS seine fishery we use harvest of hatchery pink salmon."}

d = readRDS(file=paste0("../data-generated/hatchery_diversity.rds"))
d$p  <- d$p*0.453592 / 1e6 # convert to kg

p1 = ggplot(d, aes(p, tot_div, col=year)) +
  geom_point(size=2) +
  xlab("Hatchery harvest (million kg)") +
  ylab("Catch diversity") +
  facet_wrap(~p_fshy, scale="free_x") +
  scale_color_viridis() +
  theme_sleek() + labs(colour = "Year")
p1
```

\break

```{r fig6, echo=FALSE, fig.pos="placeHere", fig.cap="Effects of the fraction of hatchery fish on (1) mean annual catches, and (2) standard deviation of annual catches, for three areas ('KO-AP' = Kodiak and Alaska Peninsula, 'PWS' = Prince William Sound, 'SE' = Southeast). Effect sizes are presented as the percentage change from an increase from 0% to 20% hatchery fish. Dots represent mean estimates and error bars represent 95% CIs."}

df = readRDS(file=paste0("../data-generated/","region","_df.rds"))

# chinook chum coho pink sock
cbPalette <- c("#009E73", "#4B2E83", "#999999", "#CC79A7", "#D55E00")

df$fishery[which(df$fishery=="Prince William Sound")] = "PWS"
df$fishery[which(df$fishery=="SEAK")] = "SE"
df$fishery[which(df$fishery=="Kodiak")] = "KO-AP"

p1 = ggplot(df, aes(x=fishery, y=bestimate, group=species, color=species)) +
  geom_point(position=position_dodge(0.5), alpha = 0.5, size=3) +
  ylab("Effect size (mean)") +
  xlab("") +
  geom_errorbar(aes(ymin=bestimate-1.96*bsd, ymax=bestimate+1.96*bsd),
    position=position_dodge(0.5), alpha=0.5) +
  scale_color_manual(values = c('Chinook'=cbPalette[1], 'Chum'=cbPalette[2],
    'Coho'=cbPalette[3], 'Pink'=cbPalette[4], 'Sockeye'=cbPalette[5]), name="Species",
    labels =c("Chinook","Chum","Coho","Pink","Sockeye"))+
  geom_hline(aes(yintercept=0), linetype = 2) +
  theme_sleek() + coord_flip() + theme(legend.position="none") + ylim(c(-0.4,0.4))

p2 = ggplot(df, aes(x=fishery, y=estimate, group=species, color=species)) +
  geom_point(position=position_dodge(0.5), alpha = 0.5, size=3) +
  ylab("Effect size (sd)") +
  xlab("") +
  geom_errorbar(aes(ymin=estimate-1.96*sd, ymax=estimate+1.96*sd),
    position=position_dodge(0.5), alpha=0.5) +
  scale_color_manual(values = c('Chinook'=cbPalette[1], 'Chum'=cbPalette[2],
    'Coho'=cbPalette[3], 'Pink'=cbPalette[4], 'Sockeye'=cbPalette[5]), name="Species",
    labels =c("Chinook","Chum","Coho","Pink","Sockeye"))+
  geom_hline(aes(yintercept=0), linetype = 2) +
  theme_sleek() + coord_flip() + ylim(c(-0.4,0.4))

plot_grid(p1, p2, align = "h", nrow = 1, rel_widths = c(0.43, 0.57))
#gridExtra::grid.arrange(p1, p2, nrow=1)
```

\break

\break


```{r printtable, echo=FALSE, results = 'asis'}
library(xtable)
m = matrix("", 6, 4)
m[1,] = c("S01A", "Purse Seine", "Southeast Alaska", "Chum, pink")
m[2,] = c("S01E", "Purse Seine", "Prince William Sound", "Chum, pink")
m[3,] = c("S01K", "Purse Seine", "Kodiak", "Pink, sockeye")
m[4,] = c("S01M", "Purse Seine", "Alaska Peninsula", "Pink, sockeye")
m[5,] = c("S03A", "Gillnet", "Southeast Alaska", "Chum, sockeye")
m[6,] = c("S03E", "Gillnet", "Prince William Sound", "Sockeye, chum, Chinook")
colnames(m) = c("Fishery (permit #)", "Gear", "Region", "Target spp. (2015)")
print(xtable(as.data.frame(m), caption="Salmon fisheries that were included in the analysis of catch diversification on revenue and revenue variability."), comment=FALSE)
```

\break

# References
